Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
update
incorporated @bpbond 's feedback on intro, fixed missing references

Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

@@

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
Yet, this sink is reduced by deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].

@@

For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., Leith, @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969 Odum, Luyssaert 2008].
- increasingly refined understanding of the elements of ecosystem C cycles,  
- standardization of definitions, (which were vague at the time of Odum)
- Expanding set of variables, notably including the development of the eddy-covariance technique to measure ecosystem-atmosphere CO~2~ exchange (REFS), NEP, soil respiration, Also various techniques for measuring belowground C cycling.
- critical review of methods and their sources of errors
- Increasing standardization of  methods: allometries, which are highly influential to any measurement including woody biomass or changes therein, but until recently have not been standardized
- Further standardization made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], GEM (Malhi), NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
- While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.

Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
In this format, the data are effectively inaccessible for many global-scale analyses, 
Data synthesis began,...

To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
*ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration). 
These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
Data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].

>>>>

Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

@@

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].

@@

For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969; @luyssaert_old-growth_2008].
Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001]. 
New technology has also enabled researches to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux (**REF-BEN**) and ecosystem-atmosphere CO~2~ exchange (**REFS**).
Standardization of estimation techniques has also increased; for example, there has been increasing standardization of the biomass allometries strongly influence estimates of any C cycle variable including woody biomass or changes therein [e.g., @chojnacky_updated_2014; @chave_improved_2014].
Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.

Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles, along with variation in data formats, units, measurement methods, *etc.*
To use these data to address global-scale questions, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020]. 
The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration). 
These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones, making it ideal for assessing how forest C cycling varies across biomes and with respect to stand age.
